---
title: "NDVI_resampling"
author: "Xavier Rotllan-Puig"
date: "`r format(Sys.time(), '%d/%m/%Y')`"
output: 
  word_document:
    #reference_docx: LPD_MS_styles.docx
    #toc: true #table of content true
    toc_depth: 3  #up to three depths of headings (specified by #, ## and ###)
    #number_sections: true  #number sections at each table header
    #theme: united  #options for theme
    #highlight: tango  #syntax highlighting style
    #css: style.css   #custom css, should be in same folder. Only for HTML
    #pandoc_args:
    #  - --lua-filter=scholar-metadata.lua
    #  - --lua-filter=author-info-blocks.lua
      
#bibliography: lpd_biblio.bib
#csl: methods-in-ecology-and-evolution.csl
always_allow_html: yes


---


```{r setup, include = FALSE, results='asis'}
library(knitr)
library(pander)
library(captioner)
knitr::opts_chunk$set(echo = TRUE)

```

```{r include = FALSE}
if(Sys.info()[4] == "D01RI1700371"){
  path2data <- "E:/rotllxa/NDVI_resample/NDVI_data"
  path2save <- "E:/rotllxa/NDVI_resample/NDVI_resample"
}else if(Sys.info()[4] == "h05-wad.ies.jrc.it"){
  path2data <- ""
  path2save <- ""
}else if(Sys.info()[4] == "MacBook-MacBook-Pro-de-Xavier.local"){
  path2data <- "/Users/xavi_rp/Documents/D6_LPD/NDVI_data"
  path2save <- "/Users/xavi_rp/Documents/D6_LPD/NDVI_resample"
}else{
  stop("Define your machine before to run LPD")
}


table_num <- captioner::captioner(prefix = "Table")
fig_num <- captioner::captioner(prefix = "Figure")
#ts <- 0
```


```{r include = FALSE}
fig0_1km <- fig_num(name = "f0_1km_300", caption = "Working NDVI maps at 1km and 333m resolution, respectively")
orig_mapsCat <- paste0(path2save, "/ndvi1km_300m_Cat.jpg")


```


This is a document to show different options for the resampling of Land products (https://land.copernicus.eu/) from 333m to 1km using R functions and packages. Some comparison of the performance of several methods are also provided.

The analisys is done using a subset of the 10-daily Normalized Difference Vegetation Index global product downloaded from the Copernicus Global Land Product portal (https://land.copernicus.eu/global/products/ndvi). The selected images are from 11^th^August 2019 and the working maps are cut between the coordinates (DD) Xmin = 0, Xmax = 4, Ymin = 40, Ymax = 43. See `r fig_num("f0_1km_300", display = "cite")` for the two working maps at 1km and 333m resolution, respectively. 


![`r fig0_1km`](`r orig_mapsCat`)







